### 2018-03-10 paper/sexSelectionSpacing-ver1.tex
Discuss middle education duration and sex ratio results

% [How to read tables]
Tables \ref{tab:median_sex_ratio_low}, \ref{tab:median_sex_ratio_med}, and
\ref{tab:median_sex_ratio_high}
show predicted median duration and sex ratios by period, spell, and sex 
composition for the three different education levels separated by area of residence.%
\footnote{
Appendix Tables \ref{tab:p25_p75_low}, \ref{tab:p25_p75_med}, and 
\ref{tab:p25_p75_low} show 25th and 75th percentiles durations.
} 
The predictions are based on the estimation method described in 
Section \ref{sec:strategy}.
The standard errors for both measures are based on bootstrapping, where the 
original sample is resampled with replacement and the method is applied to 
the new sample.

@@

The differences in duration across sex compositions are more 
pronounced with higher-order spells.
For the second spell, the difference in median duration between
having a girl or a boy as the first-born are mostly between 0.5 and 1 
month, although it varies from a low of 0.2 months for rural women 
in 1995--2004 period to a high of 1.5 months
for urban women in the 1985--1994 period.
Focusing on rural women for the third spell, the difference is,
on average, slightly over 1 months between only girls and only boys 
and the differences are statistically significant in all periods.
For rural women the differences for the fourth spell are just below
2 months, with the exception of the 1972--1984 period, which is
also the only period where the difference is not statistically significant.

@@

There are no clear time trends in the differences in median duration
across sex compositions or the predicted sex ratios across the four periods.
Despite some sex ratios that are statistically significantly higher 
than the natural sex ratio there is therefore little consistent evidence 
for the use of sex selection for women without education.
There has, for example, been relatively little change in the length
of spacing for rural women with two girls in the third spell and 
the predicted sex ratio was higher in the 1985--1994 period
than in the 2005--2016 period.
The possible exception is the fourth spell, where both duration 
and sex ratio has increased for women with three girls, although
the \emph{difference} across sex compositions has remained almost
the same.

@@

For women with between 1 and 7 years of education it is harder
to discern a pattern, partly because a larger number of area,
sex composition, and periods combinations have relatively few 
births in them.
That said, the predicted median durations show less evidence 
of son preference than for women without education.
The exception to this are for the third spell for rural women
for the 1972-1984 and 1985-1994 periods where the median
durations are two and three months shorted with only girls
compared to only boys, although the duration is longer for
only girls in the 1995-2006 period.
As for women with no education, there is little evidence of
sex selection for women with 1 to 7 years of education,
consistent with the absence of changes in relative durations
across periods.

@@

>>>>

% [How to read tables]
Tables \ref{tab:median_sex_ratio_low}, \ref{tab:median_sex_ratio_med}, and
\ref{tab:median_sex_ratio_high}
show predicted median duration and sex ratios by period, spell, and sex 
composition for the three different education levels separated by area of residence.%
\footnote{
Appendix Tables \ref{tab:p25_p75_low}, \ref{tab:p25_p75_med}, and 
\ref{tab:p25_p75_high} show 25th and 75th percentiles durations.
} 
The predictions are based on the estimation method described in 
Section \ref{sec:strategy}.
The standard errors for both measures are based on bootstrapping, where the 
original sample is resampled with replacement and the method is applied to 
the new sample.

@@

The differences in duration across sex compositions are more 
pronounced with higher-order spells.
For the second spell, the difference in median duration between
having a girl or a boy as the first-born are mostly between 0.5 and 1 
month, although it varies from a low of 0.2 months for rural women 
in 1995--2004 period to a high of 1.5 months
for urban women in the 1985--1994 period.
Focusing on rural women for the third spell, the difference is,
on average, slightly over one months between only girls and only boys 
and the differences are statistically significant in all periods.
For rural women the differences for the fourth spell are just below
2 months, with the exception of the 1972--1984 period, which is
also the only period where the difference is not statistically significant.

@@

There are no clear time trends in the differences in median duration
across sex compositions or the predicted sex ratios across the four periods.
Despite that some sex ratios are statistically significantly higher 
than the natural sex ratio there is therefore little consistent evidence 
for the use of sex selection for women without education.
There has, for example, been relatively little change in the length
of spacing for rural women with two girls in the third spell and 
the predicted sex ratio was higher in the 1985--1994 period
than in the 2005--2016 period.
The possible exception is the fourth spell, where both duration 
and sex ratio has increased for women with three girls, although
the \emph{difference} across sex compositions has remained almost
the same.

@@

Women with one to seven years of education follow a pattern
broadly similar to those with no education as 
Table \ref{tab:median_sex_ratio_med} shows.
There is a more equal distribution in the number of women
across urban and rural for this education group than for the
no education women.
As a result, the only spell where there are few women to base 
results on are for the fourth spell for urban women and the first 
period of the fourth spell for rural women.

Spacing between births increased over the four decades.
On average, the increase is approximately two months.
The exception is again the first spell, where the time from
marriage to first birth remained the same or fell slightly.

In the period before sex selection became available, 1972--1985,
the differences in spacing between only girls and only boys
increase with spell number.
For both urban and rural women the difference in duration
is just over a month for the second spell, but for the
third spell it is 1.4 and 2.3 months for urban and rural
women, respectively.
In the three later periods there is no clear pattern
in the differences across sex compositions.

The differences in duration between only girls and only boys 
appear to decrease over time.
The exception is for the two later periods where the
differences are slightly larger in the last period compared
with the second-to-last period.

The decreases in differences are accompanied by increases
in predicted sex ratios, some of which are statistically 
significantly different from the natural sex ratio.
Hence, even though the evidence is not strong, it does
appear that women with one to seven years of education 
have begun to use sex selection at higher-order spells.
This is also consistent with the falling fertility for both
urban and rural women shown below.

@@

Women with eight or more years of education appear the least likely
to follow the pattern of shorter spacing when they have only girls
compared to only boys before sex selection became available as 
Table \ref{tab:median_sex_ratio_high} shows.
Part of the problem may be that there are few women with that level
of education for the 1972--1984 period.
This means that only the results for the first and second spells in 
both urban and rural areas are likely to be reliable.
Furthermore, there are relatively few women with three boys who
go on to having a fourth birth, both in urban and rural areas.

The spacing between births increases substantially over time, especially
for urban women. 
Urban women show an increase in median duration of between a quarter
and three-quarters of a year, whereas the increase for rural women is 
mostly around a quarter of a year.
The spacing between marriage and the first birth is still the
exception, and for this group of women there is a decline of
about a month for both urban and rural areas.
